---
title: "Initial 2015-2018 Data Creation"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook contains the initial code for cleaning the 2015-2018 crime file data to create Shaw-specific shapefiles.

## Dependencies
This notebook requires:

```{r load-packages}
# primary data tools
library(compstatr)     # work with stlmpd crime data

# tidyverse packages
library(dplyr)         # data wrangling
library(readr)         # working with csv data

# spatial packages
library(sf)            # working with spatial data
library(tigris)        # tiger/line api access

# other packages
library(janitor)       # frequency tables
library(here)          # file path management
library(knitr)         # output
library(testthat)      # unit testing
```

## Create Data
Data downloaded from the STLMPD website come in `.csv` format but with the wrong file extension. The following bash script copies them to a new subdirectory and fixes the file extension issue:

```{bash}
# change working directory
cd ..
# execute cleaning script
bash source/reformatHTML.sh
```

## Load Data
With our data renamed, we build a year list objects for 2016, 2017, and 2018 crimes:

```{r load-data}
data2015 <- cs_load_year(here("data", "intermediate", "2015"))
data2016 <- cs_load_year(here("data", "intermediate", "2016"))
data2017 <- cs_load_year(here("data", "intermediate", "2017"))
data2018 <- cs_load_year(here("data", "intermediate", "2018"))
```

